Milk intake and stroke mortality in the Japan Collaborative Cohort Study - a Bayesian survival analysis

### 2020-07-17 paprfolder_docx/strokeJACC.Rmd
docx updated

Abstract:

  The aim was to further examine the relationship between milk intake and stroke mortality among the Japanese population. We used data from the Japan Collaborative Cohort Study to estimate the posterior acceleration factors (AF) as well as the hazard ratios (HR) comparing individuals with different milk intake frequencies against those who never consumed milk at the study baseline. These estimations were computed through a series of Bayesian survival models that employed a Markov Chain Monte Carlo simulation process. 100000 posterior samples for each individual were generated separately through four independent chains after model convergency were confirmed. Posterior probabilites that daily milk consumers had lower hazard or delayed mortality from strokes compared to non-consumers was 99.0% and 78.0% for men and women, respectively. Accordingly, the estimated posterior means of AF and HR for daily milk consumers were 0.88 (95% Credible Interval, CrI: 0.81, 0.96) and 0.80 (95% CrI: 0.69, 0.93) for men and 0.97 (95% CrI: 0.88, 1.10) and 0.95 (95% CrI: 0.80, 1.17) for women.

keywords: 
  milk intake; mortality; stroke; Bayesian survival anlysis; time-to-event data; JACC study

@@

Dairy food, especially milk has been recommended to reduce stroke risk by nearly 7% for each 200 g increment of daily consumption [@DeGoede2016]. More intuitive interpretation for a decreasing risk would be possible if we were able to compute the exact probability for people who had milk intake may had lower hazard of dying from stroke compared with those who never drank milk at all. For general public/media reporting, concept of hazard in epidemiological studies could still sometimes be challenging to be understood or misinterpreted since hazard is formally defined as the probability of the occurrence of an event at a given time point [@collett2015modelling]. Usually, authors of epidemiological papers would tend to use "risk" instead of "hazard" or interchangeably. However, it would still possibly be mixed up with "risk" that only contain pure meaning of "probability of an event" without redefining a point or a period of time in cross-sectional settings. For better understanding and interpretation of the findings from data that researchers endeavored to collect, statistical literature have provided plenty of choices that could help us better communicate with each other. Another approach of comparing the time-to-event survival probabilities between different groups would be to model the time before observing an event rather than the hazard which always required the assumption of a proportional hazard to be met. Accelerated failure time models are among these convenient tools that would avoid worrying about the assumption of proportional hazard and directly showing how faster/slower one individual in an exposure group might have an event compared to others among different exposure groups [@Wei1992].

Our aim was to overcome these potential pitfalls, avoid misunderstanding, and provide a more straightforward answer to the main research question that whether someone answered he/she drank milk at the baseline of study had lower hazard of dying from stroke compared with his/her counterparts who said they never consumed milk. If the answer to the primary objective was yes, then the probabilities that individuals with different frequencies of milk intake may had lower hazard compared with those who never drank milk were calculated through a Markov Chain Monte Carlo (MCMC) simulation process. A Bayesian survival analysis method was applied on an existing database and through which, we also provided estimates about whether drinking milk could delay or slow down the speed towards a mortality from stroke event from happening after controlling for the other potential confounders.

@@

We calculated sex-specific means (standard deviation, sd) and proportion of selected baseline characteristics according to the frequency of milk intake. Overall difference across the milk intake groups were tested by either analysis of variance for continuous variables or $\chi^2$ test for categorical variables.

@@

The total follow-up was 1555073 person-years (median = 19.3 years), during which 2675 death from stroke was confirmed (1352 men and 1323 women). Among these stroke mortality, 952 were hemorrhagic stroke (432 men and 520 women), and 957 were cerebral infarction (520 men and 437 women).

@@

<!-- - strength and limitation of current study -->
Some limitations here are worth mentioning. First, although our object was not to answer which type of milk is protective, but if such information were somehow available in the JACC study database, more detailed comparison or stratification would have been possible. Second, despite reasonable validity of FFQ in the JACC study cohort was assessed and confirmed, measurement errors are inevitable. Therefore, we did not try to compute the amount of consumption by multiplying an average volume per occasion with the frequency of intake since the random error might be exaggerated and the observed associations may have attenuated. Strengths of our analyses included that we have transformed the research questions to more transparent ones that is easier for interpretation. Direct probabilities that daily milk intake is associated with lower hazard or delayed stroke mortality event were provided here after thorough computer simulation.

>>>>

**Abstract**:

  The aim was to further examine the relationship between milk intake and stroke mortality among the Japanese population. We used data from the Japan Collaborative Cohort Study to estimate the posterior acceleration factors (AF) as well as the hazard ratios (HR) comparing individuals with different milk intake frequencies against those who never consumed milk at the study baseline. These estimations were computed through a series of Bayesian survival models that employed a Markov Chain Monte Carlo simulation process. 100000 posterior samples for each individual were generated separately through four independent chains after model convergency were confirmed. Posterior probabilites that daily milk consumers had lower hazard or delayed mortality from strokes compared to non-consumers was 99.0% and 78.0% for men and women, respectively. Accordingly, the estimated posterior means of AF and HR for daily milk consumers were 0.88 (95% Credible Interval, CrI: 0.81, 0.96) and 0.80 (95% CrI: 0.69, 0.93) for men and 0.97 (95% CrI: 0.88, 1.10) and 0.95 (95% CrI: 0.80, 1.17) for women. In conclusion, data from the JACC study has provided strong evidence that daily milk intake among Japanese men was associated with delayed and lower hazard of mortality from stroke especially cerebral infarction.

**keywords**: 
  milk intake; mortality; stroke; Bayesian survival anlysis; time-to-event data; JACC study

@@

Eastern Asian populations were reported to have higher burden from either mortality or morbidity from stroke than populations in European or American regions [@Kim2014]. Dairy food, especially milk has been suggested to reduce stroke risk by nearly 7% for each 200 g increment of daily consumption [@DeGoede2016]. Although 2 daily servings of milk or dairy products is recommended in Japan [@Yoshiike2007], the actual per capita intake ($\approx$ 63 g/day) of these food groups is much lower and less frequent than that in Western countries [@Saito2019]. Given that previous reports have also indicated no significant [@Iso1999;@sauvaget2003intake;@Elwood2004] or even positive associations [@Larsson2009], whether people with such a low level intake of milk can still benefit against stroke would require elucidation. A more intuitive interpretation would be available if we were able to show the probabilities and how certain the existing data can provide evidence about whether drinking milk can delay or lower the hazard of dying from stroke. Accelerated failure time models under Bayesian framework are convenient tools that would help avoid worrying about the assumption of proportional hazard normally required in a Cox proportional hazard model and can directly show how faster/slower individuals in one exposure group might have an event compared to others among different exposure groups [@Wei1992;@ibrahim2014b].

Our aim was to provide a more straightforward answer to the primary research question that whether someone answered he/she drank milk at the baseline of study had lower hazard of dying from stroke compared with his/her counterparts who said they never consumed milk. If the answer to the primary objective was yes, then the probabilities that individuals with different frequencies of milk intake may had lower hazard compared with those who never drank milk were calculated through a Markov Chain Monte Carlo (MCMC) simulation process. A Bayesian survival analysis method was applied on an existing database and through which, we also provided estimates about whether drinking milk could delay a stroke mortality event from happening after controlling for the other potential confounders.

@@

We calculated sex-specific means (standard deviation, sd) and proportion of selected baseline characteristics according to the frequency of milk intake. Age-adjusted stroke mortality rate were expressed as per 1000 person-year predicted through poisson regression models.

@@

The total follow-up was 1555073 person-years (median = 19.3 years), during which 2675 death from stroke was confirmed (1352 men and 1323 women). Among these stroke mortality, 952 were hemorrhagic stroke (432 men and 520 women), and 957 were cerebral infarction (520 men and 437 women). Age-adjusted stroke mortality rates for each category of milk intake frequency was estimated to be 1.8, 2.0, 1.7, 1.6, 1.5 per 1000 person-year for men and 1.3, 1.4, 1.2, 1.1, 1.2 per 1000 person-year for women.

@@

<!-- - strength and limitation of current study -->
Some limitations here are worth mentioning. First, the milk intake frequency as well as other lifestyle information was collected only once at the baseline and was self-reported. Apparently life habits are possible to alter over time and these would resulted in misclassfication and residual confoundings. Second, despite reasonable validity of FFQ in the JACC study cohort was assessed and confirmed, measurement errors are inevitable. Therefore, we did not try to compute the amount of consumption by multiplying an average volume per occasion with the frequency of intake since the random error might be exaggerated and the observed associations may have attenuated. Strengths of our analyses included that we have transformed the research questions to more transparent ones that is easier for interpretation. Direct probabilities that daily milk intake is associated with lower hazard or delayed stroke mortality event were provided here after thorough computer simulations.
